---
title: "ESM 204 Homework 2"
author: "Justine Lang"
date: "2023-05-01"
output: 
  html_document: 
    code_folding: hide
---

```{r setup, include=FALSE, echo = TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(tidyr)
library(kableExtra)
```

```{r}
carbon <- read_csv(here("CarbonUC.csv")) %>% 
  clean_names()
```

```{r}
###Pivot data frame longer

carbon_long <- carbon %>% 
  pivot_longer(cols = c(2:5), names_to = "campus", values_to = "mc")
```


## Question 1

For each campus, plot the marginal abatement cost data and estimate a linear regression model with zero intercept. These estimated linear regressions will be your estimates of the marginal abatement cost curve for each campus, which you will use for the remainder of this analysis. Write the equation for the marginal abatement cost curve for each campus.

```{r}
###plot abatement for ucsb

ggplot(data = carbon_long, aes(x = abatement, y = mc, color = campus)) +
  geom_point() +
  scale_color_manual(values = c('gold2','blue', 'yellow', 'navy')) +
  labs(x = "Abatement (tons C02)",
       y= "Marginal Cost of Abatement ($/ton)",
       title = "UC Carbon Abatement") +
  theme_minimal()
```



```{r}
### UCSB linear regression model 
ucsb_lm <- lm(ucsb ~ 0 + abatement, data = carbon)

### coefficient: 0.001004  

### UCLA linear regression model 
ucla_lm <- lm(ucla ~ 0 + abatement, data = carbon)

### coefficient: 0.0007014

### UCI linear regression model 
uci_lm <- lm(uci ~ 0 + abatement, data = carbon)

### coefficient: 0.002008 

### UCD linear regression model 
ucd_lm <- lm(ucd ~ 0 + abatement, data = carbon)

### coefficient: 0.003024  
```

```{r}
###slopes and baselines

ucsb_slope <- ucsb_lm$coefficient[1]
ucla_slope <- ucla_lm$coefficient[1]
uci_slope <- uci_lm$coefficient[1]
ucd_slope <- ucd_lm$coefficient[1]

ucsb_base <- 90000
ucla_base <- 250000
uci_base <- 100000
ucd_base <- 160000
```

```{r}
slope_df <- data.frame(ucsb_slope, 
                       ucla_slope,
                       uci_slope,
                       ucd_slope) %>% 
  rename(ucsb = ucsb_slope,
         ucla = ucla_slope,
         uci = uci_slope,
         ucd = ucd_slope) %>% 
  pivot_longer(ucsb:ucd,
               names_to = "campus",
               values_to = "slope")

slope_df <- slope_df %>% 
  mutate(baseline = case_when(
    campus == "ucsb" ~ ucsb_base,
    campus == "ucla" ~ ucla_base,
    campus == "uci" ~ uci_base,
    campus == "ucd" ~ ucd_base
  ))
```

```{r}
###Create predicted MC curves

ucsb_predict <- predict(ucsb_lm)
ucla_predict <- predict(ucla_lm)
uci_predict <- predict(uci_lm)
ucd_predict <- predict(ucd_lm)
```

```{r}
mc_predict_df <- data.frame(ucsb_predict, ucla_predict, uci_predict, ucd_predict) %>% 
  rename(ucsb = ucsb_predict,
         ucla = ucla_predict,
         uci = uci_predict,
         ucd = ucd_predict) %>% 
  pivot_longer(ucsb:ucd, names_to = "campus", values_to = "mc_predict")

carbon_long$mc_predict <- mc_predict_df$mc_predict
```

```{r}
###Create demand curve

carbon_long <- carbon_long %>% 
  mutate(demand = slope_df$slope*(slope_df$baseline - carbon_long$abatement))
```

```{r}
###Plot MC curves - add caption

ggplot(data = carbon_long, aes(x = abatement, color = campus)) +
  geom_line(aes(y = mc_predict), size = 0.8) + 
  scale_color_manual(values = c('gold2','blue', 'yellow', 'navy')) +
  geom_point(aes(y = mc), alpha = 0.3) +
  labs(x = '\n Tons CO2 Abated', y = 'Marginal cost of abatement ($/ton)\n', color = 'Campus') + 
  theme_minimal()
```



UCSB: $P = 0.001A$
UCLA: $P = 0.0007A$
UCI: $P = 0.002A$
UCD: $P = 0.003A$



## Question 2

Using your estimated marginal abatement cost curves, together with the baseline emissions, derive each campus’ demand curve for carbon. In other words, how much would each campus be willing to pay to emit the first ton of carbon, the second ton of carbon, etc. Plot each campus’ demand curve for emissions on a single graph. Which campus is willing to pay the most for the first ton of carbon emissions?

```{r}
###Plot demand curve - add caption

ggplot(data = carbon_long, aes(x = abatement, color = campus)) +
geom_line(aes(y = demand), size = 0.8) + 
  scale_color_manual(values = c('gold2','blue', 'yellow', 'navy')) +
  labs(x = '\n Tons CO2 Emitted', y = 'Marginal willingness to pay ($/ton)\n', color = 'Campus') + 
  theme_minimal()
```


UCD is willing to pay the most to emit the first ton of carbon. 



## Question 3

Here you will analyze three alternative interventions for reducing carbon emissions across these four campuses. Please answer the questions about each intervention listed below.

**a.** Intervention 1 - Carbon Tax: Each campus must pay a tax of $50/ton for each ton of emissions. How much would each campus choose to emit? How much would each campus abate? What is the total amount of emissions across all four campuses? How much tax revenue is generated?

```{r}
tax <- 50
```

```{r}
ggplot(data = carbon_long, aes(x = abatement, color = campus)) +
geom_line(aes(y = demand), size = 0.8) + 
  geom_line(aes( y = 50), size = 0.8) +
  scale_color_manual(values = c('gold2','blue', 'yellow', 'navy')) +
  labs(x = '\n Tons CO2 Emitted', y = 'Marginal willingness to pay ($/ton)\n', color = 'Campus') + 
  theme_minimal()
```

Come back to this.


**b.** Intervention 2 - Carbon Ban: Carbon emissions are banned at each campus. What is the total cost of this ban to each campus?

Come back to this.


**c.** Intervention 3 - Cap without Trade: Each campus is capped at 100,000 tons of carbon. How much abatement would be required by each campus to comply with this cap? What is the marginal cost of the last unit of abatement for each campus? What is the total cost of abatement for each campus to comply with this new cap?

```{r}
ucla_abate <- ucla_base - 100000
ucd_abate <- ucd_base - 100000

### UCLA would have to abate 150,000 tons. UCD would have to abate 60,000 tons
```

UCLA would have to abate 150,000 tons. UCD would have to abate 60,000 tons. UCSB and UCI do not exceed the cap and would not need to pay for abatement. 

```{r}
ucla_abate_cost <- function(a){ucla_slope * a}
ucla_abate_cost(ucla_abate)

ucd_abate_cost <- function(a){ucd_slope * a}
ucd_abate_cost(ucd_abate)
```

UCLA MAC: $105.22

UCD MAC: $181.45

```{r}
ucla_total <- (ucla_abate * ucla_abate_cost(ucla_abate))/2

ucd_total <- (ucd_abate * ucd_abate_cost(ucd_abate))/2
```

UCLA total cost: $7,891,141

UCD total cost: $5,443,456













